partial: An R Package for Creating Partial Dependence Plots

### 2016-08-10 manuscript/greenwell.tex
minor edits

The benefit of using \code{partial} is threefold: (1) it is a generic function that can be used for various types of model fits (not just random forests), (2) it will allow for any number of predictors to be used, and (3) it can utilize any of the parallel backends supported by the \CRANpkg{foreach} package \citep{foreach-pkg}; we discuss parallel execution in a later section. For example, the following code chunk could be used to assess the joint effect of \code{lstat} and \code{rm} on \code{cmedv} based on a conditional random forest using various options in \code{plotPartial}. The results are displayed in Figure~\ref{fig:pd_lstat_rm}.
\begin{example}
library(party)
boston.crf <- cforest(cmedv ~ ., data = boston)
pd.lstat.rm <- partial(boston.crf, pred.var = c("lstat", "rm"))
plotPartial(pd.lstat.rm)
plotPartial(pd.lstat.rm, col.regions = colorRampPalette(c("white", "blue")))
plotPartial(pd.lstat.rm, contour = FALSE, zlab = "cmedv", drape = TRUE,
            colorkey = TRUE, screen = list(z = -20, x = -60))
\end{example}
Note that the default color map for contour plots is the Matplotlib "viridis" color map provided through the \CRANpkg{viridis} package \citep{viridis-pkg}.

\begin{widefigure}[htbp]
  \centering
  \includegraphics[width=1.0\linewidth]{pd_lstat_rm}
  \caption{Partial dependence of \code{cmedv} on \code{lstat} and \code{rm} based on a conditional random forest. \textit{Left}: Default plot. \textit{Middle}: Using a different color palette. \textit{Right}: Using a 3-D surface.}
  \label{fig:pd_lstat_rm}
\end{widefigure}

>>>>

The benefit of using \code{partial} is threefold: (1) it is a generic function that can be used for various types of model fits (not just random forests), (2) it will allow for any number of predictors to be used, and (3) it can utilize any of the parallel backends supported by the \CRANpkg{foreach} package \citep{foreach-pkg}; we discuss parallel execution in a later section. For example, the following code chunk could be used to assess the joint effect of \code{lstat} and \code{rm} on \code{cmedv} using a multivariate adaptive regression splines (MARS) model (fit using the \pkg{earth} package) allowing for up to three-way interactions. The results, which make use of various \code{plotPartial} options, are displayed in Figure~\ref{fig:pd_lstat_rm}.
\begin{example}
library(earth)
boston.mars <- earth(cmedv ~ ., data = boston, degree = 3)
pd.lstat.rm <- partial(boston.mars, pred.var = c("lstat", "rm"))
plotPartial(pd.lstat.rm)
plotPartial(pd.lstat.rm, col.regions = colorRampPalette(c("white", "blue")))
plotPartial(pd.lstat.rm, contour = FALSE, zlab = "cmedv", drape = TRUE,
            colorkey = TRUE, screen = list(z = -20, x = -60))
\end{example}
Note that the default color map for contour plots is the Matplotlib "viridis" color map provided through the \CRANpkg{viridis} package \citep{viridis-pkg}.

\begin{widefigure}[htbp]
  \centering
  \includegraphics[width=1.0\linewidth]{pd_lstat_rm}
  \caption{Partial dependence of \code{cmedv} on \code{lstat} and \code{rm} based on a third degree MARS model. \textit{Left}: Default plot. \textit{Middle}: Using a different color palette. \textit{Right}: Using a 3-D surface.}
  \label{fig:pd_lstat_rm}
\end{widefigure}
